---
title: "P8105 Homework 2"
author: "Sarah Munro"
date: "10/7/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r eval=FALSE}
library (tidyverse)
install.packages("tidyr")
library (readxl)
```

#Problem 1

_Read and clean the Mr Trash Wheel sheet_
```{r}
mr_trash_wheel = 
  read_excel("./Data/mr_trash.xlsx", sheet = 1) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  select (-15, -16, -17) %>%
  mutate(
    sports_balls = as.integer(sports_balls)
    )
mr_trash_wheel
```
_Read and clean precipitation data for 2017_
```{r}
precip17 = read_excel("./Data/Mr_Trash_Wheel.xlsx", sheet = 6, skip = 1) %>%
  janitor::clean_names() %>%
  drop_na() %>%
  mutate(year = 2017)
precip17
```
_Read and clean precipitation data for 2018_
```{r}
precip18 = read_excel("./Data/Mr_Trash_Wheel.xlsx", sheet = 5, skip = 1) %>%
  janitor::clean_names() %>%
  drop_na() %>%
  mutate(year = 2018)
precip18
```
_Merge datasets_
```{r}
precip =
  full_join(precip17, precip18) %>%
  mutate(month = as.character(month))
precip
```
_Write a paragraph about the datasets with inline r code_
The dataset `mr_trash_wheel` has `r nrow(mr_trash_wheel)` observations and the merged precipitation dataset `precip` has `r nrow(precip)` observations. Key variables include `dumpster`and `total` precipitation. The median of `sports_balls` is `r median(pull(filter(mr_trash_wheel, year == 2017)sports_balls))`. The total precipitation in 2018 was `r sum(pull(precip18, total))`. 

#Problem 2

_Clean the data in pols-month_
```{r}
politics =
  read_csv(file = "./Data/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day")) %>%
mutate(
  year = as.numeric(year),
  month = as.numeric(month),
  month = month.abb[month],
  day = as.numeric(day)
) %>%
mutate(
  president = if_else(prez_gop == 1, "gop", "dem")
) %>%
 select(-day, -prez_gop, -prez_dem)
politics
```

_Clean the data in snp.csv_
```{r}
snp =
  read_csv(file = "./Data/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, into = c("month", "day", "year")) %>%
  arrange(year, month) %>%
mutate(
  year = as.numeric(year),
  month = as.numeric(month),
  month = month.abb[month]
) %>%
  select(year, month, -day, close) 
snp
```
_Clean the data in unemployment.csv_
```{r}
unemployment =
  read_csv(file ="./Data/unemployment.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "rate" 
  ) 
unemployment
```
_Join the datasets_
```{r}
politics_snp = 
left_join(politics, snp, by = c("year", "month"))
final = 
left_join(politics_snp, unemployment, by = c("year", "month"))
final
```

The dataset `snp` has `r nrow(snp)` observations. The dataset `politics` has `r nrow(politics)` observations. The dataset `unemployment` has `r nrow(unemployment)` observations. The joined dataset `final` has `r nrow(final)` observations. This data combines some key variables from the other dataset such as `president`, `close`, and `rate`. It covers the range of years `r range(pull(final, year))`. 

#Problem 3 

library(dplyr)
library(stringr)

_Tidy the baby names dataset_
```{r}
baby_names =
  read_csv("./Data/baby_names.csv") %>%
  janitor::clean_names() %>%
  distinct() %>%
  mutate(
    childs_first_name = str_to_upper(childs_first_name) 
  ) %>%
  mutate(
    ethnicity = recode(ethnicity, "BLACK NON HISP" = "BLACK NON HISPANIC", "WHITE NON HISP" = "WHITE NON HISPANIC", "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER")
  )
baby_names
```
_Produce a table of the rank of the name Olivia over time_
```{r}
baby_names %>%
  filter (childs_first_name == "OLIVIA") %>%
  select (-count) %>%
  pivot_wider(names_from = year_of_birth,
              values_from = rank) %>%
knitr::kable()
```
_Produce a table of the top ranking male names over time_
```{r}
baby_names %>%
  filter (gender == "MALE", rank ==1) %>%
  select (-count) %>%
  pivot_wider(names_from = year_of_birth,
              values_from = rank) %>%
knitr::kable()
```
_Produce a scatterplot of rank of white non hispanic male names in 2016 vs the count of those names_
```{r}
scatterplot = 
  filter (baby_names, gender =="MALE", ethnicity =="WHITE NON HISPANIC", year_of_birth ==2016) 
  ggplot(scatterplot, aes(x = rank, y = count)) + geom_point()

```



  